\documentclass[12pt]{sides}

% Packages
\usepackage{fontspec}
\usepackage{xunicode}
\usepackage{multicol}
\usepackage{xspace}
\usepackage[margin=20pt]{geometry}
\usepackage{everypage}
% TODO: 
% Page border
% Fancy initials
% 'fi' ligature in scena secunda, Cap. speaking last line
% Italic 'C' made taller/longer



\setmainfont[Numbers=OldStyle,Ligatures=Discretionary,Style=Historic,StylisticSet=18]{Junicode}
\setlength\columnseprule{.3pt}
\AddEverypageHook{}

% Character Commands
\newcommand{\s}{{\fontspec[Style=]{Junicode}s}}

% Drama commands


% Document
\begin{document}
	\title{THE TRAGEDIE OF MACBETH}
	\author{William Shakespeare}
	\date{Latter Month{\s} of 1623}
	\maketitle
	
	\castpagelatin
	\cast{Macbeth}{}
	\cast{Lady Macbeth}{}
	\cast{Banquoh}{}
	\cast{Fleance}{}
	\cast{King Malcolme}{}
	\cast{Donalbaine}{}
	\cast{Macduff}{}
	\cast{Lady Macduff}{}
	\cast{Macduff's son}{}
	\cast{Lenox}{}
	\cast{Ross}{}
	\cast{Menteith}{}
	\cast{Angus}{}
	\cast{Caithness}{}
	\cast{Siward}{}
	\cast{Young Siward}{}
	\cast{Three Witches}{}
	
	\newpage
	
	\newactnamed{Actu{\s} Primu{\s}.}

	\begin{multicols}{2}		
		\newscenenamed{Sc{\oe}na Prima}
			\stagedir{Thunder and lightning. Enter three Witche{\s}.}
			
			\dia{\oldstylenums{1}.} When shall we three meet againe? \\ In Thunder,\allowbreak Lightning,\allowbreak or in Raine?
			
			\dia{\oldstylenums{2}.} When the Hurley-Burley'{\s} done, \\ When the Battaile'{\s} lost,and wonne.
			
			\dia{\oldstylenums{3}.} That will be ere the set of Sunne.
			
			\dia{\oldstylenums{1}.} Where the place?
			
			\dia{\oldstylenums{2}.} Vpon the Heath,
			
			\dia{\oldstylenums{3}.} There to meet with \textit{Macbeth}.
			
			\dia{\oldstylenums{1}.} I come,\textit{Gray-Malkin}.
			
			\dia{All.} \textit{Padock} call{\s} anon: faire i{\s} foule,and foule i{\s} faire, \\ Houer through the fogge and filthie ayre.
			\stagedir{Exeunt}
			
		\newscenenamed{Scena Secunda}
			\stagedir{Alarum within. Enter King Malcome,\allowbreak Donalbaine,\allowbreak Lenox,\allowbreak with attendant{\s},meeting a bleeding Captaine.}
			\dia{King.} What bloody man i{\s} that? he can report, \\  A{\s} seemth by hi{\s} plight,\allowbreak of the Reuolt \\ The newest state.
			
			\dia{Cap.} Doubtfull it stood,\\ a{\s} two spent Swimmer{\s},that doe cling together,\\ And choake their Art: The mercilesse \textit{Macdonwald} \\ (Worthie to be a Rebell,for to that \\ The multiplying Villanie{\s} of Nature \\ Doe swarme vpon him) from the Westerne Isle{\s} \\ Of Kerne{\s} and Gallowgrosse{\s} i{\s} supply'd, \\ And Fortune on hi{\s} damned Quarry smiling, \\ Shew'd like a Rebell{\s} Whore: but all'{\s} too weake: \\ For braue \textit{Macbeth}( well hee deserue{\s} that Name) \\ Disdayning Fortune,with hi{\s} brandisht Steele, \\ Which smoak'd with bloody execution \\ (Like Valour{\s} Minion) caru'd out hi{\s} passage, \\ Till hee fac'd the Slaue: \\ Which neu'r shooke hand{\s},\allowbreak nor bad farwell to him, \\ Till he vnseam'd him from the Naue toth Chop{\s}, \\ And {fi}x'd hi{\s} Head vpon our Battlement{\s}.
			
			\dia{King.} O valiant Cousin,\allowbreak worthy Gentleman.
			
			\dia{Cap.} A{\s} whence the Sunne gin{\s} hi{\s} reflection, \\ Shipwracking Storme{\s},\allowbreak and direfull Thunder{\s}: \\ So from that Spring,\allowbreak whence comfort seem'd to come, \\ Discomfort swell{\s}: Marke King of Scotland,\allowbreak marke, \\ No sooner Iustice had,\allowbreak with Valour arm'd, \\ Compell'd these {\s}kipping Kerne{\s} to trust their heele{\s}, \\ but the Norweyan Lord,\allowbreak surueying vantage, \\ With furbusht Arme{\s},\allowbreak and new supplye{\s} of men, \\ Began a fresh assault.
			
			\dia{King.} Dismay'd not thi{\s} our Captaine{\s},\allowbreak \textit{Macbeth} and \textit{Banquoh}?
			
			\dia{Cap.} Ye{\s},\allowbreak a{\s} Sparrow{\s},\allowbreak Eagle{\s}; \\ Or the Hare, the Lyon: \\ If I say sooth, I must report they were \\ A{\s} Cannon{\s} ouer-charg'd with double Crack{\s}, \\ So they doubly redoubled stroake{\s} vpon the Foe: \\ Except they meant to bathe in reeking Wound{\s}, \\ Or memorize another Golgotha, \\ I cannot tell: but I am faint, \\ My gashe{\s} cry for helpe.
			
			\dia{King.} So well thy word{\s} become thee,\allowbreak a{\s} thy wound{\s}, \\ They smack of Honor both: Goe get him Surgeon{\s}.
			
			\stagedir{Enter Rosse and Angu{\s}.}
			Who come{\s} here?
			
			\dia{Mal.} The worthy \textit{Thane} of Rosse.
			 
			\dia{Lenox.} What a haste looke{\s} through his eye{\s}? \\ So should he looke, that seeme{\s} to speake thing{\s} strange.
			
			\dia{Rosse.} God saue the King.
			
			\dia{King.} Whence cam'st thou,\allowbreak worthy \textit{Thane} ?
			
			\dia{Rosse.} From Fiffe,\allowbreak great King, \\ Where Norweyan Banner{\s} flowt the Skie, \\ And fanne our people cold. \\ \textit{Norway} himselfe,\allowbreak with terrible number{\s} \\ Assisted by that most disloyall Traytor, \\ The \textit{Thane} of Cawdor,\allowbreak began a dismall Conflict, \\ Till that \textit{Bellona'{\s}} Bridegroome,\allowbreak lapt in proofe, \\ Confronted him with selfe-comparison{\s}, \\ Point against Point,\allowbreak rebelliou{\s} Arme 'gainst Arme, \\ Curbing hi{\s} lauish spirit: and to conclude, \\ The Victorie fell on v{\s}.
			
			\dia{King.} Great Hapinesse.
			
			\dia{Rosse.} That now,\allowbreak \textit{Sweno},\allowbreak the Norwaye{\s} King, \\ Craue{\s} composition: \\ Nor would we deigne him buriall of hi{\s} men, \\ Till he di{\s}bursed,\allowbreak at Saint \textit{Colme{\s}} ynch, \\ Ten thousand Dollar{\s},\allowbreak to our generall vse.
			
			\dia{King.} No more that \textit{Thane} of Cawdor shall decieue \\ Our bosome interest: Goe pronounce hi{\s} present death, \\ And with hi{\s} former Title greet \textit{Macbeth}.
			
			\dia{Rosse.} Ile see it done.
			
			\dia{King.} What he hath lost,\allowbreak Noble \textit{Macbeth} hath wonne.
			\stagedir{Exeunt.}
			
		\newscenenamed{Scena Tertia.}	
			\stagedir{Thunder. Enter the three Witches.}
			
			\dia{1.} Where hast thou beene,\allowbreak Sister?
			
			\dia{2.} Killing Swine.
			
			\dia{3.} Sister,where thou?
			
			\dia{1.} A Saylor{\s} Wife had Chestnut{\s} in her Lappe, \\ And mouncht,\allowbreak \& mouncht,\allowbreak and mouncht: \\ giue me, quoth I. \\ Aroynt thee,\allowbreak Witch,\allowbreak the rumpe-fed Ronyon Crye{\s}. \\ Her Husband'{\s} to Aleppo gone,\allowbreak Master o'th'\textit{Tiger:} \\ But in a Syue Ile thither sayle, \\ And like a Rat without a tayle, \\ Ile doe,\allowbreak Ile doe,\allowbreak and Ile doe.
			
			\dia{2.} Ile giue thee a Winde.
			
			\dia{1.} Th'art kinde.
			
			\dia{3.} And I another.
			
			\dia{1.} I my selfe haue all the other, \\ And the very Port{\s} they blow, \\ All the Quarter{\s} that they know, \\ I'th'Ship-man{\s} Card. \\  Ile dreyne him drie as Hay: \\ Sleepe shall neyther Night nor Day \\ Hang vpon his Pent-house Lid:			
	\end{multicols}
\end{document}